---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv('activity.csv')
data$date <- as.Date(data$date, "%Y-%m-%d")
```


## What is mean total number of steps taken per day?
Histogram showing frequencies of sums of steps taken per day
```{r}
hist(with(data, tapply(steps, date, FUN=sum)))
```
Mean of sums of steps taken per day 
```{r}
mean(with(data, tapply(steps, date, FUN=sum)), na.rm=T)
```  
Median of sums of steps taken per day
```{r}
median(with(data, tapply(steps, date, FUN=sum)), na.rm=T)
```

## What is the average daily activity pattern?
```{r}
meanSteps <- with(data, tapply(steps, interval, FUN=mean, na.rm=T))
plot(meanSteps, type='l')

meanSteps[meanSteps==max(meanSteps)]
```


## Imputing missing values
The amount of missing vaules:
```{r}
sum(is.na(data$steps))
```
Create new dataset with missing values filled with mean for a given interval:
```{r}
datb <- data
datb <- within(datb, steps <- ifelse(!is.na(steps), steps, meanSteps))
```
## What is mean total number of steps taken per day?
Histogram showing frequencies of sums of steps taken per day after substituting missing values:
```{r}
hist(with(datb, tapply(steps, date, FUN=sum)))
```
Mean of sums of steps taken per day 
```{r}
mean(with(datb, tapply(steps, date, FUN=sum)))
```  
Median of sums of steps taken per day
```{r}
median(with(datb, tapply(steps, date, FUN=sum)))
```

## Are there differences in activity patterns between weekdays and weekends?
Function asking whether it is weekday or not:
```{r}
isWeekday <- function(date) {
	if(weekdays(date)=="Saturday"|weekdays(date)=="Sunday") {
	form <- 'weekend'}
	else {form <- 'weekday'}
	return(form)
}
```
Add a column 'weekpart' using isWeekday:
```{r}
weekpart <- as.character(lapply(datb$date, isWeekday))
datb <- cbind(datb, weekpart=weekpart)
head(datb)
```
Split marked data:
```{r}
s <- split(datb, datb$weekpart)
```
Plot:
```{r}
par(mfrow = c(2, 1))
plot(with(s$weekday, tapply(steps, interval, FUN=mean)), main="Weekdays", type="l")
plot(with(s$weekend, tapply(steps, interval, FUN=mean)), main="Weekends", type="l")
```
